### 2012-10-24 paper_sean/results.tex
changes in results before Jeff's read - swapped figure positions: fitness and servo pos - changed figures for fitness and frequency

\edit{Punt for now, update this once we have results from Kyrre} 
The gaits evolved with HyperNEAT in simulation
significantly outperformed the gaits evolved with a GA by Glette et
al. 2012~(Table~\ref{results}). We compared Kyrre et al 2012's
best simulated run to the average of our 20 runs and found that our
gait is, in average, greater in fitness by 7.67 cm/s, with 2.0 cm/s
standard deviation, for every generation of evolution during the
runs. Using the same data, we recorded a fitness of 21.0 cm/s with
4.29 cm/s standard deviation as the mean fitness throughout the
evolutionary process while Kyrre et al 2012's data recorded a fitness
of 13.35 cm/s with 4.06 cm/s standard deviation. These results show
that in simulation, HyperNEAT's gaits evolve faster and have higher
fitnesses than those of GA's.

Also add: This result
is even more significant when one considers that the gait produced by
HyperNEAT with a simulator evolved with 20000 (33\%) fewer evaluations
per run than Glette et al. 2012, a decision mandated by the available
computational resources.

@@

\begin{table}
\begin{center}
\begin{tabular}{|l|r|r|r|r|}
\hline
                                         &              & Simulated  & Real Vel. & Real Vel.  \\
                                         & Evaluations  & Velocity &    (CCML)     & (ROBIN) \\
\hline
Parameterized gaits + optimization \cite{yos:clune}   &153    & --    & 5.8 & --\\
\hline
HyperNEAT in hardware \cite{yos:clune}                 & 180         & --         &   9.7  & --   \\
\hline
RL PoWER Spline \cite{haocheng}                         & 300         & --         &   11.1 & --\\
\hline
GA + simulator \cite{glette}             & 60000       & 16.4       &   13.0   & \textbf{17.8}  \\
\hline
HyperNEAT + simulator [this paper]                     & 40000       & \textbf{25.4}       &   \textbf{14.5} & --\\
\hline
\end{tabular}
\end{center}
\tablabel{results}
\caption{Gait speeds in simulation and on each of two QuadraTot hardware platforms for this paper and three previous studies. We report the number of evaluations per run, simulated (average) and real trial (max) velocities in cm/s.}  %\tablabel{results}
\end{table}

Qualitatively, the HyperNEAT gaits are regular and coordinated, as has
been found in previous simulation
studies~\cite{clune2009evolving,clune2011performance}. This is evident
in the servo position plots where the gaits produced in this study are
smooth and are similar between the servos~(see \figref{servo_plot_111}
for a representative example). This result is important because it
confirms that HyperNEAT can produce the important property of
regularity in simulation, which was not previously observed when
evolving directly on the hardware~\cite{yos:clune}.

@@

\figgp{avg_fit_vs_gen}{.48}{avg_freq_vs_gen}{.48}
{Performance plots. \textbf{Left: }Fitnesses of the champion gaits
averaged over 20 runs, measured in distance travelled over 12 seconds
in cm/s. These gaits outperformed gaits evolved in simulation using a
genetic algorithm~\cite{glette} by 54.3\% in simulation and 5.4\% in
reality, and they outperformed gaits evolved using HyperNEAT in
hardware by 49.5\% (\tabref{results}). \edit{Add Kyrre's runs (+ Haocheng?), switch
from STD to SE if necessary.}\textbf{Right: }Average frequencies of
the gaits averaged over 20 runs. After implementing a policy which
punished gaits with high frequencies to reduce servo shutdowns,
HyperNEAT was able to find high-performing, low frequency
gaits. \edit{add line at 1.67 Hz}}

\figp{servo_plot_111}{0.95}
{Servo positions over time (for nine servos) for a gait produced in
this study. The smoothness, symmetry, and regularities of the gait
show that by using a simulator, HyperNEAT was able to fully exploit
the geometry of the problem.}

>>>>

A simulated physical environment provides a noise-free setting for
testing learning algorithms. The resulting gaits in simulation after
learning with HyperNEAT reveal that HyperNEAT is capable of evolving
high fitness and highly organic-looking gaits when many evolutionary
evaluations can be afforded.

We compared the average fitnesses of 20 learning runs performed in
simulation using HyperNEAT and a GA\footnote{For a more
robust comparison to our 20 runs, we ran 19 more runs with the same
learning settings as used in Kyrre et al. 2012~\cite{glette}, in
addition to the 1 run published.}, with settings used in Glette et
al. 2012~\cite{glette} and found that HyperNEAT outperformed the
latter: $\emph{p} < 6.8\cdot10^{-8}$, comparing the
20 learning runs from both learning methods.\footnote{\emph{p} value generated
with MATLAB's non-parametric Mann-Whitney \emph{U}-test} 
HyperNEAT's gaits are in average of
55\% faster at the final learning generation --- HyperNEAT: mean 25.4
cm/s $\pm$ 3.4 SD, and Genetic Algorithm: mean 16.7 cm/s $\pm$ 1.9 SD
(\tabref{results}).

%Also add: This result
%is even more significant when one considers that the gait produced by
%HyperNEAT with a simulator evolved with 20000 (33\%) fewer evaluations
%per run than Glette et al. 2012, a decision mandated by the available
%computational resources.

@@

\begin{table}
\begin{center}
\begin{tabular}{|l|r|r|r|r|}
\hline
                                         &              & Simulated  & Real Vel. & Real Vel.  \\
                                         & Evaluations  & Velocity &    (CCML)     & (ROBIN) \\
\hline
Parameterized gaits + optimization \cite{yos:clune}   &153    & --    & 5.8 & --\\
\hline
HyperNEAT in hardware \cite{yos:clune}                 & 180         & --         &   9.7  & --   \\
\hline
RL PoWER Spline \cite{haocheng}                         & 300         & --         &   11.1 & --\\
\hline
GA + simulator \cite{glette}             & 60000       & *16.7       &   13.0   & \textbf{17.8}  \\
\hline
HyperNEAT + simulator [this paper]                     & 40000       & \textbf{25.4}       &   \textbf{14.5} & --\\
\hline
\end{tabular}
\end{center}
\tablabel{results}
\caption{Gait speeds in simulation and on each of two QuadraTot hardware platforms for this paper and three previous studies. We report the number of evaluations per run, simulated (average) and real trial (max) velocities in cm/s. *Instead of using the single fitness value reported in~\cite{glette}, we ran 19 additional runs and used the average fitness of those runs.}  %\tablabel{results}
\end{table}

Qualitatively, the HyperNEAT gaits are regular and coordinated, as has
been found in previous simulation
studies~\cite{clune2009evolving,clune2011performance}. This is evident
in the servo position plots where the gaits produced in this study are
smooth and are similar between the servos~(see \figref{servo_plot_111}
for a representative example). This result is important because it
confirms that HyperNEAT can produce the important property of
regularity in simulation, which was not previously observed when
evolving directly on the hardware~\cite{yos:clune}. HyperNEAT's higher
performance, as well as its higher standard deviation, is likely due
to HyperNEAT exploring and exploiting a greater variation of gaits,
some of which are better than others.

@@

\figp{final_sean_kyrre_fitness}{0.92}
{Fitnesses of the champion gaits evolved with HyperNEAT and a genetic algorithm
averaged over 20 runs, measured in distance travelled over 12 seconds
in cm/s. The dashed lines are standard deviations. HyperNEAT gaits outperformed gaits evolved in simulation using a
genetic algorithm~\cite{glette} by 54.3\% in simulation and 5.4\% in
reality, and they outperformed gaits evolved using HyperNEAT in
hardware by 49.5\% (\tabref{results}).}

\figgp{servo_plot_111}{.48}{frequencyPlotThreshFinal}{.48}
{\textbf{Left: }
Servo positions over time (for nine servos) for a gait produced in
this study. The smoothness, symmetry, and regularities of the gait
show that by using a simulator, HyperNEAT was able to fully exploit
the geometry of the problem.
\textbf{Right: }Average frequencies of the gaits averaged over 20 runs. 
After implementing a policy which
punished gaits with high frequencies to reduce servo shutdowns,
HyperNEAT was able to find high-performing, low frequency
gaits.}
